Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-15 ieee_paper.tex
typo
section 4 minor

We discuss how the implementation of the v4m4 kernel works using
vectorial instruction. (We will show that an implementation relying on
compiler vectorization does not lead to desirable performance.) The
registers in the MIC architecture are 512 bits wide and can store 16
floats or integers. (In comparison to OpenCL or CUDA, one can think of
these registers as a warp.) The end goal is to be able to load and
format the non zero values $a,b,c,d$ and the vector entries $p,q,r,s$
in two vectorial register to be able to apply Fused Multiply-Add on
them, which is depicted in Figure~\ref{fig:tensor_product}. Bringing
the data in the vector register as depicted in the figure is the
difficult part that we now explain.

@@

The core of the technique is the use of swizzling and permutation
features of the vectors in the Intel MIC architecture. A vector of 512
bits is composed of 4 lanes of 128 bits which are made of 4 elements
of 32 bits. A swizzling operation allows to reorder the elements
within each lane. While a permutation allow to change the orders of
the lanes. Figure~\ref{fig:swizzling_permutation} illustrates these
two possibilities. Both permutation and swizzling support common
reordering and broadcast.

@@

  \caption{Swizzling enables to reorder or broadcast elements within
    each 128 bit group. Permutation enables to reorder or broadcast
    while 128 bit groups.}
  \label{fig:swizzling_permutation}
\end{figure}

>>>>

We discuss how the implementation of the v4m4 kernel works using
vectorial instruction. (We will show that an implementation relying on
compiler vectorization does not lead to desirable performance.) The
registers in the MIC architecture are 512 bits wide and can store 16
floats or integers. (In comparison to OpenCL or CUDA, one can think of
these registers as a warp.) The end goal is to be able to load and
format the non zero values $a,b,c,d$ and the vector entries $p,q,r,s$
in two vectorial register to be able to apply Fused Multiply-Add on
them, which is depicted in Figure~\ref{fig:tensor_product}. Bringing
the data in the vector register as depicted in the figure is the
difficult part that we now explain. A thread will perform the
multiplications one row at a time, and parallelism is achieved by
giving blocks of rows to each core usign an OpenMP construct.

@@

The core of the technique is the use of swizzling and permutation
features of the vectors in the Intel MIC architecture. A vector of 512
bits is composed of 4 lanes (sometimes called channels) of 128 bits
which are made of 4 elements of 32 bits. A swizzling operation allows
to reorder the elements within each lane. While a permutation allow to
change the orders of the lanes. Figure~\ref{fig:swizzling_permutation}
illustrates these two possibilities. Both permutation and swizzling
support common reordering and broadcast.

@@

  \caption{Swizzling enables to reorder or broadcast elements within
    each 128 bit group. Permutation enables to reorder or broadcast
    whole 128 bit groups.}
  \label{fig:swizzling_permutation}
\end{figure}

All the operations that loads data from the memory into a register
allows to load 512 bits, so 16 floats at a time. A single non zero in
the matrix will cause 16 operations to be performed, but it only uses
4 floats from the matrix and 4 floats from the {\tt x} vector. So we
need to load at once the data for 4 non zeroes in the matrix, so that
we load 16 flaots from the matrix and 16 floats from the vectors at
once. (The number of non zeroes on a row in our application is always
a multiple of 4, if it was, we would need to add some explicit zeroes
in the data structure to pad it.)
